---
title: "Product Analysis using R"
output: rmarkdown::github_document
always_allow_html: true
# output: 
#   html_notebook:
#     toc: true
#     toc_float:
#       collapsed: false
#       smooth_scroll: true
#     number_sections: false
#     theme: lumen

---    

```{r library, message=FALSE, include=TRUE}
# Libraries

# Wrangling
library(tidyverse)
library(lubridate)
library(scales)

# Summary Statistics
library(outliers)

# Text Analysis
library(stringr)        # text cleaning and regular expressions
library(tidytext)       # provides additional text mining functions
library(widyr)

# Clustering
library(factoextra)
library(cluster)

# Assoc Rules
library("arules")
library("arulesViz")

# Plotting
library(ggplot2)
library(ggrepel)
library(gridExtra)
library(GGally)
library(ggridges)
library(kableExtra)
library(wesanderson)

# Mapping
library(leaflet)
library(sf)
library(RColorBrewer)

# Network Visualization
library(igraph)
library(ggraph)
library(visNetwork)
library(data.table)

```

```{r parameters, message=FALSE, include=TRUE}
# Input Path
file_orders <- here::here("03 data/orders.rds")
file_products <- here::here("03 data/product.rds")
file_customers <- here::here("03 data/customers_gender_zip_geocode.rds")
file_co <- here::here("03 data/customers_gender_zip_geocode_orders.rds")
file_zip_ph <- here::here("03 data/zip_ph.rds")

# Read Files
orders.raw <- read_rds(file_orders)
products.raw <- read_rds(file_products)
customers.raw <- read_rds(file_customers)
co.raw <- read_rds(file_co)

# Number Formatting
options(scipen = 999)

# Parameters
# Irrelevant customer_ids
xtest <- c("3273557147747", "3297613348963", "3436247875683", "3210539368547", 
           "4412203910", "2236354953315")

# Table Preferences for Kable
tb.ops <- c("striped", "hover", "condensed", "responsive")

```

```{r message=FALSE, include=TRUE}
# Customer Product Data 
products.data <- 
  products.raw %>% 
  inner_join(orders.raw, by = c("order_number" = "order_number")) %>%
  filter(order_year >= 2019) %>%
  rename(customer_id = customer_id.x) %>%
  mutate(has_value = if_else(item_value > 0, "has value", "zero value"),
         has_customer = if_else(is.na(customer_id), "no customer id", "has customer id"))

```


### Product Analysis Overview

&nbsp;  
This analysis provides an overview of the website's product assortment focusing on specialty hair products. 

Products were categorize according to  
  1. **price**  
  2. **brand**  
  3. **primary need** (e.g., coloring, thinning hair, or damaged hair)  
  
  **product type** (e.g., shampoo, conditioner, or treatment) could no be determined for 30% of the orders.  
  As a result, it is excluded from the analysis.



### Using text-analysis to group products
The network diagram below shows the most common **product need** and **product type** for each brand, based on the descriptors used in the website. Each brand's respective websites were also consulted to confirm its positioning.

```{r products.visnet.data, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# PART 3: Customers and Product Keywords

# select correlation parameters
cor.max = 1.1
cor.min = 0.5
count.min = 5

#TF IDF Analysis
words.data <- 
   products.data %>%
   select(product_id, product_features, item_value) %>%
   filter(!is.na(product_features), !is.na(product_id)) %>% 
   select(product_id, product_features) %>%
   distinct() %>%
   unnest_tokens(word, product_features)



words.data <- 
        words.data %>%
        count(product_id, word, sort = TRUE) %>%
        ungroup()

words.data <- 
    words.data %>%
        bind_tf_idf(word, product_id, n)

word.list <-
  words.data %>%
  group_by(product_id) %>%
  arrange(desc(tf_idf)) %>%
  mutate(word_rank = row_number()) %>%
  filter(word_rank <= 2) %>%
  ungroup() %>%
  select(word) %>%
  distinct()

brand.list <-
  products.data %>% select(product_brand1) %>% na.omit() %>% distinct() %>% pull(product_brand1)

# format word.data
words.data <- 
  products.data %>%
  select(product_features) %>% distinct() %>% mutate(id = row_number()) %>%
  filter(!is.na(product_features)) %>%
  unnest_tokens(word, product_features) %>%
  inner_join(word.list) %>%
  group_by(word) %>% filter(!word == "azta") %>%
  filter(n() >= 10)  

# transform words.data into a network list
network.data <- 
  words.data %>%
  # pairwise_cor(word, customer_id, upper = FALSE) %>%
  # filter(!is.na(correlation),
  #        correlation < cor.max, 
  #        correlation > cor.min)
  ungroup() %>%
  pairwise_count(word, id, upper = FALSE, sort = TRUE) %>% 
  filter(!is.na(n),
         #n < cor.max, 
         n > count.min)

## prepare node.size
# node.size <-
#   words.data %>%
#   count("id" = word, wt = item_value/100000, name = "size")

# Calculate degrees
degree_value <- 
  network.data %>%
  graph_from_data_frame() %>%
  degree()

node.size <- 
  degree_value %>% as_tibble(rownames = "id") %>% rename("size" = "value") %>%
  mutate(size = if_else(id %in% brand.list, 5, size * 5))


# VisNetwork: interactive network graph  
vis.data <- 
  network.data %>%
  graph_from_data_frame() %>%
  toVisNetworkData()

# modify nodes and edges
vis.data$nodes <- 
  vis.data$nodes %>%
  mutate(font.size = 200) %>%
  inner_join(node.size, by = c("id" = "id"))


# vis.data$edges <- vis.data$edges %>% mutate(width = correlation*10, length = 300/correlation)
vis.data$edges <- vis.data$edges %>% 
  mutate(length = case_when(from %in% brand.list ~ 10,
                            to %in% brand.list ~ 10,
                            TRUE ~0)) %>% 
  filter(length > 0) %>%
  filter(!(from %in% brand.list & to %in% brand.list))
  #mutate(width = n/10,length = 10000/n)

# complile graph

vis <-
  visNetwork(nodes = vis.data$nodes, edges = vis.data$edges) %>%
  visPhysics(solver = "forceAtlas2Based",  
             forceAtlas2Based = list(gravitationalConstant = -1000)) %>%
  visLayout(randomSeed = 123) %>%
  visOptions( highlightNearest = T )

```

```{r products.visnet, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
vis
```

### Pricing distribution
The ridge plots below show how certain product groups are typically priced higher, or lower than the average. It is also evident some product groups, such as *colored hair* and *L'oreal* occupy a wide range of price points. 

```{r products.price.plot.data, message=FALSE, warning=FALSE, include=FALSE}
## PLOT PREPARATION: PRODUCT PRICE ANALYSIS
# geom_ridge 
products.data2 <-
products.data[rep(seq_along(products.data$item_qty), products.data$item_qty), ]

# Brand Price Distribution
p1 <-  
  products.data2 %>%
  filter(!is.na(product_brand1)) %>%
  group_by(product_brand1) %>%
  mutate(levels = median(item_price_net)) %>%
  ggplot(aes(y = reorder(product_brand1,levels), 
             x = item_price_net, 
             alpha = 0.1, 
             fill = levels)) +
  # geom_boxplot() +
  geom_density_ridges() +
  scale_y_discrete(name = NULL) +
  scale_x_log10(name = NULL,
                n.breaks = 6,
                labels = scales::dollar_format(prefix = "₱")) +
  # scale_fill_discrete( name = NULL, guide = FALSE) +
  scale_fill_gradientn(name = NULL, guide = FALSE, colours = wes_palette("Zissou1", n = 17, type = "continuous")) +
  scale_size(guide = NULL) +
  scale_alpha(guide = NULL) +
  labs(title = "Price Distribution\nby Brand") +
  theme(
    plot.margin = unit(c(1,1.2,0.5,0.5),"cm"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    axis.text.x = element_text(angle = 45, hjust = 1))  

# Need Price Distribution
p2 <-  
  products.data2 %>%
  filter(!is.na(product_need1),!product_need1 %in% c("salon")) %>%
  group_by(product_need1) %>%
  mutate(levels = median(item_price_net)) %>%
  ggplot(aes(y = reorder(product_need1,levels), 
             x = item_price_net, 
             alpha = 0.1, 
             fill = levels)) +
  # geom_boxplot() +
  ggridges::geom_density_ridges() +
  scale_y_discrete( name = NULL) +
  scale_x_log10(name = NULL,
                n.breaks = 6,
                labels = scales::dollar_format(prefix = "₱")) +
  scale_fill_gradientn(name = NULL, guide = FALSE, colours = wes_palette("Cavalcanti1", n = 17, type = "continuous")) +
  scale_size(guide = NULL) +
  scale_alpha(guide = NULL) +
  labs(title = "Price Distribution\nby Primary Need") +
  theme(
    plot.margin = unit(c(1,1.2,0.5,0.5),"cm"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    axis.text.x = element_text(angle = 45, hjust = 1))


# Type Price Distribution
p3 <-  
  products.data2 %>%
  filter(!is.na(product_type1)) %>%
  group_by(product_type1) %>%
  mutate(levels = median(item_price_net)) %>%
  ggplot(aes(y = reorder(product_type1,levels), 
             x = item_price_net, 
             alpha = 0.1, 
             fill = product_type1)) +
  ggridges::geom_density_ridges() +
  scale_y_discrete( name = NULL) +
  scale_x_log10(name = NULL,
                n.breaks = 6,
                labels = scales::dollar_format(prefix = "₱")) +
  scale_fill_discrete( name = NULL, guide = FALSE) +
  scale_size(guide = NULL) +
  scale_alpha(guide = NULL) +
  labs(title = "Price Distribution by Type") +
  theme(
    plot.margin = unit(c(1,1.2,0.5,0.5),"cm"),
    legend.position = "bottom",
    legend.text = element_text(size = 8),
    axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r products.price.plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
## PRODUCT PRICE
grid.arrange(p2, p1, ncol = 2, nrow = 1)
```  

### Product positioning
Brands can also be plotted against sales volume and average price to determine their relative market position. In the chart below, size and opacity indicate the brand's share of total company revenue.

```{r brand.sales.plot.data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
## PLOT PREPARATION: Brand Price Volume Analysis

# Product x Customer Brand
plot.data <- products.data %>%
  filter(!is.na(product_brand1)) %>%
  group_by(product_brand1, product_need1) %>%
  summarise(item_value = sum(item_value),
            item_sales = sum(item_qty)) %>%
  group_by(product_brand1) %>%
  mutate(ylevels = sum(item_value)) %>%
  group_by(product_need1) %>%
  mutate(xlevels = sum(item_value))

p1 <-
  plot.data %>%
  group_by(product_brand1) %>%
  summarise(item_value = sum(item_value),
            item_sales = sum(item_sales)) %>%
  mutate(avg_price = item_value / item_sales) %>%
  ggplot(aes(y = avg_price,
             x = item_sales,
             # size = log(item_value),
             ), 
         ) +
  geom_point(aes(size = item_value, color = "red")) +
  geom_text_repel(aes(label = product_brand1, alpha = item_value)) +
  scale_y_continuous(name = "Average Price\n", labels = comma) +
  # scale_y_log10(name = "average price", labels = comma) +
  scale_x_continuous(name = "Total Items" ,labels = comma, limits = c(-1000,12500)) +
  # scale_x_log10(name = "items sold" ,labels = comma,  limits = c(1,1e5)) +
  scale_color_discrete(guide = NULL) +
  scale_size(guide = NULL) +
  scale_alpha(guide = NULL) +
  coord_cartesian(xlim = c(-1000,12500), ylim = c(0,4000)) +
  labs(title = "Brand Sales \nby Volume and Price") +
  theme(plot.margin = unit(c(1,1.2,0.5,0.5),"cm"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r brand.sales.plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
## PLOT: Brand Price Volume Analysis 1
p1
```


Plotting the median line shows each each brand's impact on the portfolio. Davines has the highest volume, and sells at the average price. Kerastase, Nioxin and Olaplex are high-price products, but Keratase also sells high-volumes.  Meanwhile, L'oreal and Tigi sell similar volumes, but Tigi's selling price is below average.

```{r brand.sales.plot2.data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
## PLOT: Brand Price Volume Analysis 2
p2 <-
  p1 + 
  geom_hline(yintercept = 1710, color = "black", linetype = (1), alpha = 0.8) +
  geom_rect(aes(xmin = -1000, xmax = 2500, ymin = 0, ymax = 2000), 
            color = "red", alpha = 0, linetype = (3)) +
  geom_rect(aes(xmin = -1000, xmax = 2500, ymin = 2200, ymax = 4000), 
            color = "Dark Green", alpha = 0, linetype = (3)) +
  geom_rect(aes(xmin = 6250, xmax = 8500, ymin = 2700, ymax = 4000), 
            color = "blue", alpha = 0, linetype = (3)) +
  geom_rect(aes(xmin = 9500, xmax = 11400, ymin = 1000, ymax = 2500), 
            color = "purple", alpha = 0, linetype = (3)) + 
  geom_rect(aes(xmin = 4100, xmax = 6400, ymin = 1800, ymax = 2100), 
            color = "orange", alpha = 0, linetype = (3)) +
  geom_rect(aes(xmin = 4100, xmax = 6400, ymin = 900, ymax = 1600), 
            color = "brown", alpha = 0, linetype = (3))
```

```{r brand.sales.plot2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
p2
```


Plotting the brands against product needs shows how some top brands, such as Nioxin and Olaplex, are highly focused on specific needs, while Kerastase, davines, L'oreal, and TIGI cater to a wider set of needs.


```{r brand.need.plot.data, message=FALSE, warning=FALSE, include=FALSE, paged.print=TRUE}
## PLOT PREPARATION: Brand x Need Analysis
p1 <-
  plot.data %>% filter(!is.na(product_need1)) %>%
  ggplot(aes(x = reorder(product_need1, desc(xlevels)),
             y = reorder(product_brand1, ylevels),
             color = product_need1, alpha = 0.2)) +
  geom_point(aes(size = item_sales)) +
  scale_x_discrete(name = NULL) +
  scale_y_discrete(name = NULL) +
  scale_color_discrete(guide = NULL) +
  scale_size(guide = NULL) +
  scale_alpha(guide = NULL) + 
  labs(title = "Brand Sales \nby Customer Need") +
  theme(plot.margin = unit(c(1,1.2,0.5,0.5),"cm"),
        axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r brand.need.plot, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE}
## PLOT: Brand x Need Strategy
p1
```



